A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2021-01-21 cusp_j/paper/submit/paper.tex
minor modifs from Julien

Even if these schemes are extremely efficient to obtain a qualitative description of the wave function through a meaningful selection in an exponentially growing Hilbert space, a quantitative description is often out of reach as it implies the use of very large one-electron basis sets which are needed to recover the slowly converging correlation effects near the electron-electron coalescence point. 
These short-range correlation effects are mainly due to the diverging character of the coulomb potential, which induces a non analytical behaviour of the wave function at small inter-electronic distance ($r_{12}\approx 0$): the famous electronic cusp originally derived by Kato\cite{Kat-CPAM-57}. For a clear and general derivation of the cusp conditions up to second order for Coulombic systems, see Ref. \onlinecite{Tew-JCP-08}. As shown by the seminal works of Hylleraas\cite{Hyl-ZP-29} and further developed byKutzelnigg and coworkers, \cite{Kut-TCA-85,KutKlo-JCP-91, NogKut-JCP-94} the slow convergence of the correlation energy with respect to the quality of the basis set is mainly due to the impossibility to represent the cusp in a finite one-electron basis set. 
As the cusp conditions originate from the divergence of the Coulomb potential at $r_{12}=0$, an alternative approach would be to develop theories dealing with a smooth potential. Such a path was followed by three distinct branches: 
the explicitly correlated methods (F12), transcorrelated methods (TC) and the range-separated density functional theory (RS-DFT).

In RS-DFT\cite{Sav-INC-96a,TouColSav-PRA-04}, the electron-electron interaction is split into a smooth long-range part and a complementary short-range diverging part, the former being treated explicitly by a wave function and the latter by a density functional. 
Such a splitting is done through the function $\text{erf}(\mu r_{12})$ where $\mu$ is the range separation parameter which allows one to continually move from Kohn-Sham DFT ($\mu=0$) to pure WFT ($\mu=\infty$). 
As the effective Hamiltonian obtained in RS-DFT is smooth at $r_{12}=0$ for a finite value of $\mu$, the corresponding eigenfunction has no electron-electron cusp\cite{GorSav-PRA-06} and therefore the convergence of the results with respect to the basis set are exponential\cite{FraMusLupTou-JCP-15}. 
A number of approximate RS-DFT schemes have been developed involving single-reference\cite{AngGerSavTou-PRA-05, GolWerSto-PCCP-05, TouGerJanSavAng-PRL-09,JanHenScu-JCP-09,TouZhuSavJanAng-JCP-11, MusReiAngTou-JCP-15,KalTou-JCP-18,KalMusTou-JCP-19} and multi-reference\cite{LeiStoWerSav-CPL-97, FroTouJen-JCP-07, FroCimJen-PRA-10, HedKneKieJenRei-JCP-15, HedTouJen-JCP-18, FerGinTou-JCP-18} WFT methods. 
Nevertheless, there are still some open issues in RSDFT, such as remaining fractional-charge and fractional-spin errors in the short-range density functionals \cite{MusTou-MP-17} or the dependence of the quality of the results on the value of the range-separation parameter $\mu$.

@@

An alternative point of view using a Jastrow factor is the so-called transcorrelated (TC) approach where the full effect of the Jastrow factor is incorporated into the calculation through a similarity transformation of the original Hamiltonian by the Jastrow factor. 
Seminal equations were derived by Hirschfelder\cite{Hirschfelder-JCP-63} who obtained an effective non hermitian operator based on a specific form of Jastrow factor, and later on Boys, Handy and co-workers\cite{BoyHan-PRSLA-69,BoyHanLin-1-PRSLA-69,BoyHanLin-2-PRSLA-69} derived the equations of the TC Hamiltonian for a general exponential form of the Jastrow factor which insures size extensivity. 
The obtained TC Hamiltonian contains certain new features with respect to the usual Hamiltonian: the TC Hamiltonian contains an additional effective two- and three-electron scalar potential together with is a non hermitian two-electron differential operator. Because of its non-hermitian nature, the TC Hamiltonian looses the variational principle, and Boys and Handy managed to derive equations to optimize both orbitals and Jastrow parameters for a single Slater determinant. 
Nevertheless, such a constrained form of the wave function for the Slater together with the loose of variational principle makes it difficult to find a monotonic convergence of computed energies\cite{Handy-MolPhys-71}.

@@

In the present work, we derive a new form of Jastrow factor such that the leading $1/r_{12}$ terms of the corresponding TC Hamiltonian reproduce the RS-DFT effective Hamiltonian. 
As in RS-DFT, the new TC Hamiltonian and Jastrow factor are tuned by a unique range separation parameter $\mu$. 
The aim of this paper is to establish the analytical form of the TC Hamiltonian and perform a numerical study on the ground state eigenvalues and eigenvectors on a set of two-electron atomic and molecular systems: the helium isoelectronic series from H$^-$ to Ne$^{8+}$ and H$_2$ molecule. 
In the context of the TC theory the interesting features of the present work are that 
i) the analytical form of the TC Hamiltonian is explicitly known and it turns out that all two-electron integrals can be performed analytically and the three-electron integrals can be efficiently computed in mixed numerical-analytical way, ii) the new Jastrow factor and corresponding TC Hamiltonian are tuned by a unique parameter $\mu$ which allows some flexibility, iii) there always exists a regime of $\mu$ which significantly improve the basis set convergence of WFT, iv) we propose schemes to find a value of $\mu$ which automatically adapts to the system and improves basis set convergence.

@@

To better understand the physical content of at $r_{12}\approx 0$,  
let us Taylor expand the function $u(r_{12};\mu)$ 
\begin{equation}
 \label{eq:j_dl}
 u(r_{12};\mu) = -\frac{1}{2\sqrt{\pi}\mu} + \frac{1}{2}r_{12} - \frac{\mu}{2\sqrt{\pi}} r_{12}^2 + o(r_{12}^4).
\end{equation}
A first remark is that $u(r_{12};\mu)$ exactly restores the cusp as
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0} \deriv{}{r_{12}}{}u(r_{12};\mu) = \frac{1}{2},
\end{equation}
implying that the right-eigenvectors of the TC Hamiltonian obtained with $u(r_{12};\mu)$ are necessary cusp less (see Appendix \ref{sec:cusp}). 
The Taylor expansion of Eq. \eqref{eq:j_dl} inserted in the exponential correlating Jastrow factor reads 
\begin{equation}
 \label{eq:bigJ_dl}
 e^{u(r_{12},\mu)}\approx e^{-\frac{1}{2\sqrt{\pi}\mu}} e^{ \frac{1}{2}r_{12}} e^{- \frac{\mu}{2\sqrt{\pi}} r_{12}^2}. 
\end{equation}
From Eq. \eqref{eq:bigJ_dl} one can analyze each term: i) as $e^{-\frac{1}{2\sqrt{\pi}\mu}}<1$, the Jastrow factor digs the Coulomb hole, 
ii) the term in $e^{ \frac{1}{2}r_{12}} $ exactly imposes the cusp; iii) and the last term $e^{- \frac{\mu}{2\sqrt{\pi}} r_{12}^2}$ reshapes the curvature of the Coulomb hole.

@@

Eq. \eqref{ht_def_g} leads to the following transcorrelated Hamiltonian 
\begin{equation}
 \begin{aligned}
 \label{ht_def_gu}
 \tilde{H}[\mu] & = H - \sum_{i<j} \hat{K}_\mu(\bri{i},\bri{j}) - \sum_{i<j<k} \hat{L}_\mu(\bri{i},\bri{j},\bri{k}),
 \end{aligned}
\end{equation}
where the effective two- and three-body operators $\hat{K}_\mu(\bri{1},\bri{2})$ and $\hat{L}_\mu(\bri{1},\bri{2},\bri{3})$ are defined as
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{K}_\mu(\bri{1},\bri{2})  = \frac{1}{2} \bigg( &\Delta_1 u(r_{12},\mu) + \Delta_2 u(r_{12},\mu) \\
                                               + &\big(\nabla_1 u(r_{12},\mu) \big) ^2 + \big(\nabla_1 u(r_{12},\mu) \big) ^2 \bigg) \\
                                               + &\nabla_1 u(r_{12},\mu) \cdot \nabla_2 + \nabla_2 u(r_{12},\mu)\cdot \nabla_1,
 \end{aligned}
\end{equation}
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{L}_\mu(\bri{1},\bri{2},\bri{3})  = &   \nabla_1 u(r_{12},\mu) \cdot \nabla_1 u(r_{13},\mu) \\
                                          + & \nabla_2 u(r_{21},\mu) \cdot \nabla_2 u(r_{23},\mu)  \\
                                          + & \nabla_3 u(r_{13},\mu) \cdot \nabla_3 u(r_{32},\mu).
 \end{aligned}
\end{equation}

As shown in the Appendix \ref{k_l_appendix}, the operator $\hat{K}_\mu(\bri{1},\bri{2})$ has a straightforward analytical expression which reads 
\begin{equation}
 \label{eq:k_final}
  \begin{aligned}
   \hat{K}_\mu(\bri{1},\bri{2}) = & \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} - \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2} + \frac{\bigg(1 -     \text{erf}(\mu r_{12}) \bigg)^2}{4} \\
& - \bigg( \text{erf}(\mu r_{12}) - 1\bigg) \deriv{}{r_{12}}{},
  \end{aligned}
\end{equation}
and similarly the analytical expression of $\hat{L}_\mu(\bri{1},\bri{2},\bri{3}) $ reads 
\begin{equation}
 \label{eq:l_final}
 \begin{aligned}
 \hat{L}_\mu(\bri{1},\bri{2},\bri{3}) = & \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \bri{12} \cdot \frac{1 - \text{erf}(\mu r_{13})}{2 r_{13}} \bri{13} \\
                                      + & \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \bri{21} \cdot \frac{1 - \text{erf}(\mu r_{23})}{2 r_{23}} \bri{23} \\
                                      + & \frac{1 - \text{erf}(\mu r_{13})}{2 r_{13}} \bri{31} \cdot \frac{1 - \text{erf}(\mu r_{32})}{2 r_{32}} \bri{32}.
 \end{aligned}                    
\end{equation}

By injecting the form of Eqs. \eqref{eq:k_final} and \eqref{eq:l_final} in Eq. \eqref{ht_def_gu}, one obtains 
\begin{equation}
  \begin{aligned}
 \label{eq:exp_ht_mu}
   \tilde{H}[\mu] = & h_c + \sum_{i<j} \bigg( \tilde{\mathcal{W}}_{ee}(r_{ij},\mu) + \tilde{t}[\mu] \bigg) - \sum_{i<j<k} \hat{L}_\mu(\bri{1},\bri{2},\bri{3})
  \end{aligned}
\end{equation}
where $ \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)$ is the scalar effective two-body potential of  $\tilde{H}[\mu]$
\begin{equation}
 \begin{aligned}
 \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)  =  \frac{\text{erf}(\mu r_{ij})}{r_{ij}} + \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{ij} \big)^2} - \frac{\bigg(1 -     \text{erf}(\mu r_{ij}) \bigg)^2}{4}, 
 \end{aligned}
\end{equation}
and $\tilde{t}[\mu]$ is a non hermitian two-body differential operator 
\begin{equation}
 \begin{aligned}
 \tilde{t}[\mu] =  \bigg( \text{erf}(\mu r_{ij}) - 1\bigg) \deriv{}{r_{ij}}{}.
 \end{aligned}
\end{equation}

@@

In order to find a systematic way to determine a reasonable value of $\mu$, one must keep in mind the physical meaning of such quantity: it has the unit of the inverse of a distance, and determines, at leading order, the depth of the hole imposed by the Jastrow factor. 
Therefore, it must have the typical scale of the inverse of correlation effects, which of course strongly depends on the system through its density for instance. In the context of RS-DFT, Toulouse \textit{et. al.}\cite{TouColSav-JCP-05} have investigated different flavour of $\mu$ varying in space through the density of the system in a given point in space. 
More specifically, they introduced (see Eq. (12) of \onlinecite{TouColSav-JCP-05}) a range separation parameter typical for correlation effects in the uniform electron gas (UEG) 
\begin{equation}
 \mursc({\bf r}) =  \frac{2\sqrt{\alpha / \pi}}{\sqrt{r_s({\bf r})}},
\end{equation}
with $\alpha = (9 \pi/4)^{-1/3}$, $r_s = (n 4 \pi /3)^{-1/3}$ and $n$ being the density. 
Such function $\mursc({\bf r})$ depends on the system and on the position in space through the density $n({\bf r})$ of the system at a given point ${\bf r}$. 
Therefore, we propose here to use the average value of $\mursc({\bf r})$ over the Hartree Fock electronic density to define the value of $\mu$ for a specific system in a given basis set:
\begin{equation}
 \label{eq:mu_av_rsc}
 \murscav = \frac{1}{N_e}\int \text{d}{\bf r} \mursc({\bf r}) n_{\text{HF}}({\bf r}) 
\end{equation}
where $n_{\text{HF}}({\bf r})$ is the HF density and $N_e$ the number of electrons in the system. 
We report in Table \ref{table_conv_e_mu_iso} the convergence with the basis set $\basis$ of the error with respect to the exact ground state energies of the isoelectronic series of the helium atom using $\murscav$ in $\tilde{H}[\mu]$. 
We also report in Table \ref{table_conv_e_mu_iso} the value of $\murscav$ in the double-zeta quality basis for each system, the value in larger basis sets varying by less than $0.1\%$. 
As one can observe from  Table \ref{table_conv_e_mu_iso}, the ground state energies of $\tilde{H}[\murscav]$ converge much faster than that of the usual Hamiltonian as the MAD for the ground state energies is of 2.64 mH, 0.48 mH and 0.26 mH in the double-, triple- and quadruple zeta quality basis sets (respectively) whereas it is of 12.16 mH, 4.71 mH and 1.64 mH for the usual Hamiltonian. Also, one can observe that, except for the low density systems such as the H${^-}$ and He species, it converges from below the exact ground state energy. Regarding now the value of $\murscav $, one can notice that it increases with the atomic number. This is expected as the density becomes more picked near the nucleus while the nuclear charge increases, which induces necessary that the typical inter electronic distance lowers with $Z$.

@@

From the Figs \ref{fig:H2}, one can observe that i) all methods provide a near chemical accuracy from the AVTZ basis set, ii) the TC Hamiltonian obtained with the FROGG and that with $\mfrogg$ provide very similar results, iii) the values obtained with $\muuegav$ and $\murscav$ provide very similar results and improve accuracy over the FROGG for the AVDZ basis set, iv) a value of $\mu = 0.5$ provides a good description in the AVDZ basis set while remaining above the exact energy in the AVTZ basis set, which is not the case for $\mu=0.3$.

@@

\subsection{Analytical form of $\hat{L}(\bri{1},\bri{2},\bri{3})$ and computation of related integrals }

@@

 &  \matelem{\phi_i \phi_j \phi_m}{\hat{L}(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} = \\
 +& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} \\
 +& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_2(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} \\
 +& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_3(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} 

@@

 \hat{L}_1(\bri{1},\bri{2},\bri{3}) = w_\mu(\bri{12}) w_\mu(\bri{13}) \bri{12} \cdot \bri{13}

@@

 & \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} = \\

@@

 &  \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n}  \\
 =& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1^x(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} \\  
+ & \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1^y(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} \\  
+ & \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1^z(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n},  

@@

& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1^x(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} \\  

@@

Therefore, the matrix element of the $\hat{L}_1(\bri{1},\bri{2},\bri{3})$ is simply 

@@

 & \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} \\

@@

 & \matelem{\phi_i \phi_j \phi_m}{\hat{L}(\bri{1},\bri{2},\bri{3})}{\phi_k \phi_l \phi_n} \\

>>>>

Even if these schemes are extremely efficient to obtain a qualitative description of the wave function through a meaningful selection in an exponentially growing Hilbert space, a quantitative description is often out of reach as it implies the use of very large one-electron basis sets which are needed to recover the slowly converging correlation effects near the electron-electron coalescence point. 
These short-range correlation effects are mainly due to the diverging character of the coulomb potential, which induces a non analytical behaviour of the wave function at small inter-electronic distance ($r_{12}\approx 0$): the famous electronic cusp originally derived by Kato\cite{Kat-CPAM-57}. For a clear and general derivation of the cusp conditions up to second order for Coulombic systems, see Ref. \onlinecite{Tew-JCP-08}. As shown by the seminal works of Hylleraas\cite{Hyl-ZP-29} and further developed by Kutzelnigg and coworkers, \cite{Kut-TCA-85,KutKlo-JCP-91, NogKut-JCP-94} the slow convergence of the correlation energy with respect to the quality of the basis set is mainly due to the impossibility to represent the cusp in a finite one-electron basis set. 
As the cusp conditions originate from the divergence of the Coulomb potential at $r_{12}=0$, an alternative approach would be to develop theories dealing with a smooth potential. Such a path was followed by three distinct branches: 
the explicitly correlated methods (F12), transcorrelated methods (TC) and the range-separated density functional theory (RS-DFT).

In RS-DFT\cite{Sav-INC-95,TouColSav-PRA-04}, the electron-electron interaction is split into a smooth long-range part and a complementary short-range diverging part, the former being treated explicitly by a wave function and the latter by a density functional. 
Such a splitting is done through the function $\text{erf}(\mu r_{12})$ where $\mu$ is the range separation parameter which allows one to continually move from Kohn-Sham DFT ($\mu=0$) to pure WFT ($\mu=\infty$). 
As the effective Hamiltonian obtained in RS-DFT is smooth at $r_{12}=0$ for a finite value of $\mu$, the corresponding eigenfunction has no electron-electron cusp\cite{GorSav-PRA-06} and therefore the convergence of the results with respect to the basis set are exponential\cite{FraMusLupTou-JCP-15}. 
A number of approximate RS-DFT schemes have been developed involving single-reference\cite{AngGerSavTou-PRA-05, GolWerSto-PCCP-05, TouGerJanSavAng-PRL-09,JanHenScu-JCP-09,TouZhuSavJanAng-JCP-11, MusReiAngTou-JCP-15,KalTou-JCP-18,KalMusTou-JCP-19} and multi-reference\cite{LeiStoWerSav-CPL-97, FroTouJen-JCP-07, FroCimJen-PRA-10, HedKneKieJenRei-JCP-15, HedTouJen-JCP-18, FerGinTou-JCP-18} WFT methods. 
Nevertheless, there are still some open issues in RSDFT, such as remaining fractional-charge and fractional-spin errors in the short-range density functionals \cite{MusTou-MP-17} or the dependence of the quality of the results on the value of the range-separation parameter $\mu$.

@@

An alternative point of view using a Jastrow factor is the so-called transcorrelated (TC) approach where the full effect of the Jastrow factor is incorporated into the calculation through a similarity transformation of the original Hamiltonian by the Jastrow factor. 
Seminal equations were derived by Hirschfelder\cite{Hirschfelder-JCP-63} who obtained an effective non hermitian operator based on a specific form of Jastrow factor, and later on Boys, Handy and co-workers\cite{BoyHan-PRSLA-69,BoyHanLin-1-PRSLA-69,BoyHanLin-2-PRSLA-69} derived the equations of the TC Hamiltonian for a general exponential form of the Jastrow factor which insures size extensivity. 
The obtained TC Hamiltonian contains certain new features with respect to the usual Hamiltonian: the TC Hamiltonian contains an additional effective two- and three-electron scalar potential together with is a non hermitian two-electron differential operator. Because of its non-hermitian nature, the TC Hamiltonian looses the variational principle, and Boys and Handy managed to derive equations to optimize both orbitals and Jastrow parameters for a single Slater determinant. 
Nevertheless, such a constrained form for the Slater part of the wave function together with the loose of variational principle makes it difficult to find a monotonic convergence of computed energies\cite{Handy-MolPhys-71}.

@@

In the present work, we derive a new form of Jastrow factor such that the leading $1/r_{12}$ terms of the corresponding TC Hamiltonian reproduce the RS-DFT effective Hamiltonian. 
As in RS-DFT, the new TC Hamiltonian and Jastrow factor are tuned by a unique range separation parameter $\mu$. 
The aim of this paper is to establish the analytical form of the TC Hamiltonian and perform a numerical study on the ground state eigenvalues and eigenvectors on a set of two-electron atomic and molecular systems: the helium isoelectronic series from H$^-$ to Ne$^{8+}$ and H$_2$ molecule. 
In the context of the TC theory the interesting features of the present work are that 
i) the analytical form of the TC Hamiltonian is explicitly known and it turns out that all two-electron integrals can be performed analytically and the three-electron integrals can be efficiently computed in mixed numerical-analytical way, ii) the new Jastrow factor and corresponding TC Hamiltonian are tuned by a unique parameter $\mu$ which allows some flexibility, iii) there always exists a regime of $\mu$ which significantly improves the basis set convergence of WFT, iv) we propose schemes to find a value of $\mu$ which automatically adapts to the system and improves basis set convergence.

@@

To better understand the physical content of at $r_{12}\approx 0$,  
let us Taylor expand the function $u(r_{12};\mu)$ 
\begin{equation}
 \label{eq:j_dl}
 u(r_{12};\mu) = -\frac{1}{2\sqrt{\pi}\mu} + \frac{1}{2}r_{12} - \frac{\mu}{2\sqrt{\pi}} r_{12}^2 + O(r_{12}^4).
\end{equation}
A first remark is that $u(r_{12};\mu)$ exactly restores the cusp as
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0} \deriv{}{r_{12}}{}u(r_{12};\mu) = \frac{1}{2},
\end{equation}
implying that the right-eigenvectors of the TC Hamiltonian obtained with $u(r_{12};\mu)$ are necessary cusp less (see Appendix \ref{sec:cusp}). 
The Taylor expansion of Eq. \eqref{eq:j_dl} inserted in the exponential correlating Jastrow factor reads 
\begin{equation}
 \label{eq:bigJ_dl}
 e^{u(r_{12},\mu)}\approx e^{-\frac{1}{2\sqrt{\pi}\mu}} e^{ \frac{1}{2}r_{12}} e^{- \frac{\mu}{2\sqrt{\pi}} r_{12}^2}. 
\end{equation}
From Eq. \eqref{eq:bigJ_dl} one can analyze each term: i) as $e^{-\frac{1}{2\sqrt{\pi}\mu}}<1$, the Jastrow factor digs the Coulomb hole, 
ii) the term in $e^{ \frac{1}{2}r_{12}} $ exactly imposes the cusp; iii) and the last term $e^{- \frac{\mu}{2\sqrt{\pi}} r_{12}^2}$ reshapes the curvature of the Coulomb hole.

@@

Eq. \eqref{ht_def_g} leads to the following transcorrelated Hamiltonian 
\begin{equation}
 \begin{aligned}
 \label{ht_def_gu}
 \tilde{H}[\mu] & = H - \sum_{i<j} \hat{K}(\bri{i},\bri{j},\mu) - \sum_{i<j<k} \hat{L}(\bri{i},\bri{j},\bri{k},\mu),
 \end{aligned}
\end{equation}
where the effective two- and three-body operators $\hat{K}(\bri{1},\bri{2},\mu)$ and $\hat{L}(\bri{1},\bri{2},\bri{3},\mu)$ are defined as
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{K}(\bri{1},\bri{2},\mu)  = \frac{1}{2} \bigg( &\Delta_1 u(r_{12},\mu) + \Delta_2 u(r_{12},\mu) \\
                                               + &\big(\nabla_1 u(r_{12},\mu) \big) ^2 + \big(\nabla_1 u(r_{12},\mu) \big) ^2 \bigg) \\
                                               + &\nabla_1 u(r_{12},\mu) \cdot \nabla_2 + \nabla_2 u(r_{12},\mu)\cdot \nabla_1,
 \end{aligned}
\end{equation}
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{L}(\bri{1},\bri{2},\bri{3},\mu)  = &   \nabla_1 u(r_{12},\mu) \cdot \nabla_1 u(r_{13},\mu) \\
                                          + & \nabla_2 u(r_{21},\mu) \cdot \nabla_2 u(r_{23},\mu)  \\
                                          + & \nabla_3 u(r_{13},\mu) \cdot \nabla_3 u(r_{32},\mu).
 \end{aligned}
\end{equation}

As shown in the Appendix \ref{k_l_appendix}, the operator $\hat{K}(\bri{1},\bri{2},\mu)$ has a straightforward analytical expression which reads 
\begin{equation}
 \label{eq:k_final}
  \begin{aligned}
   \hat{K}(\bri{1},\bri{2},\mu) = & \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} - \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2} + \frac{\bigg(1 -     \text{erf}(\mu r_{12}) \bigg)^2}{4} \\
& - \bigg( \text{erf}(\mu r_{12}) - 1\bigg) \deriv{}{r_{12}}{},
  \end{aligned}
\end{equation}
and similarly the analytical expression of $\hat{L}(\bri{1},\bri{2},\bri{3},\mu) $ reads 
\begin{equation}
 \label{eq:l_final}
 \begin{aligned}
 \hat{L}(\bri{1},\bri{2},\bri{3},\mu) = & \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \bri{12} \cdot \frac{1 - \text{erf}(\mu r_{13})}{2 r_{13}} \bri{13} \\
                                      + & \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \bri{21} \cdot \frac{1 - \text{erf}(\mu r_{23})}{2 r_{23}} \bri{23} \\
                                      + & \frac{1 - \text{erf}(\mu r_{13})}{2 r_{13}} \bri{31} \cdot \frac{1 - \text{erf}(\mu r_{32})}{2 r_{32}} \bri{32}.
 \end{aligned}                    
\end{equation}

By injecting the form of Eqs. \eqref{eq:k_final} and \eqref{eq:l_final} in Eq. \eqref{ht_def_gu}, one obtains 
\begin{equation}
  \begin{aligned}
 \label{eq:exp_ht_mu}
   \tilde{H}[\mu] = & h_c + \sum_{i<j} \bigg( \tilde{\mathcal{W}}_{ee}(r_{ij},\mu) + \tilde{t}[\mu] \bigg) - \sum_{i<j<k} \hat{L}(\bri{1},\bri{2},\bri{3},\mu)
  \end{aligned}
\end{equation}
where $ \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)$ is the scalar effective two-body potential of  $\tilde{H}[\mu]$
\begin{equation}
 \begin{aligned}
 \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)  =  \frac{\text{erf}(\mu r_{ij})}{r_{ij}} + \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{ij} \big)^2} - \frac{\bigg(1 -     \text{erf}(\mu r_{ij}) \bigg)^2}{4}, 
 \end{aligned}
\end{equation}
and $\tilde{t}[\mu]$ is a non hermitian two-body differential operator 
\begin{equation}
 \begin{aligned}
 \tilde{t}[\mu] =  \bigg( \text{erf}(\mu r_{ij}) - 1\bigg) \deriv{}{r_{ij}}{}.
 \end{aligned}
\end{equation}

@@

In order to find a systematic way to determine a reasonable value of $\mu$, one must keep in mind the physical meaning of such quantity: it has the unit of the inverse of a distance, and determines, at leading order, the depth of the hole imposed by the Jastrow factor. 
Therefore, it must have the typical scale of the inverse of correlation effects, which of course strongly depends on the system through its density for instance. In the context of RS-DFT, Toulouse \textit{et. al.}\cite{TouColSav-JCP-05} have investigated different flavour of $\mu$ varying in space through the density of the system in a given point in space. 
More specifically, they introduced (see Eq. (12) of \onlinecite{TouColSav-JCP-05}) a range separation parameter typical for correlation effects in the uniform electron gas (UEG) 
\begin{equation}
 \mursc({\bf r}) =  \frac{2\sqrt{\alpha / \pi}}{\sqrt{r_s({\bf r})}},
\end{equation}
with $\alpha = (9 \pi/4)^{-1/3}$, $r_s = (n 4 \pi /3)^{-1/3}$ and $n$ being the density. 
Such function $\mursc({\bf r})$ depends on the system and on the position in space through the density $n({\bf r})$ of the system at a given point ${\bf r}$. 
Therefore, we propose here to use the average value of $\mursc({\bf r})$ over the Hartree Fock electronic density to define the value of $\mu$ for a specific system in a given basis set:
\begin{equation}
 \label{eq:mu_av_rsc}
 \murscav = \frac{1}{N_e}\int \text{d}{\bf r} \mursc({\bf r}) n_{\text{HF}}({\bf r}) 
\end{equation}
where $n_{\text{HF}}({\bf r})$ is the HF density and $N_e$ the number of electrons in the system. 
We report in Table \ref{table_conv_e_mu_iso} the convergence with the basis set $\basis$ of the error with respect to the exact ground state energies of the isoelectronic series of the helium atom using $\murscav$ in $\tilde{H}[\mu]$. 
We also report in Table \ref{table_conv_e_mu_iso} the value of $\murscav$ in the double-zeta quality basis for each system, the value in larger basis sets varying by less than $0.1\%$. 
As one can observe from  Table \ref{table_conv_e_mu_iso}, the ground state energies of $\tilde{H}[\murscav]$ converge much faster than that of the usual Hamiltonian as the MAD for the ground state energies is of 2.64 mH, 0.48 mH and 0.26 mH in the double-, triple- and quadruple zeta quality basis sets (respectively) whereas it is of 12.16 mH, 4.71 mH and 1.64 mH for the usual Hamiltonian. Also, one can observe that, except for the low density systems such as the H${^-}$ and He species, it converges from below the exact ground state energy. Regarding now the value of $\murscav $, one can notice that it increases with the atomic number. This is expected as the density becomes more picked near the nucleus while the nuclear charge increases, which induces necessary that the typical inter electronic distance lowers with $Z$. In a similar fashion, we tried another definition of $\mu$ based on the work of Toulouse \textit{et. al.}\cite{TouColSav-JCP-05} which was proportional to $\frac{1}{r_s({\bf r})}$ and found that it leads to values of $\mu$ that are too large and therefore do not lead to significant improvement of the basis set convergence.

@@

From the Figs \ref{fig:H2}, one can observe that i) all methods provide a near chemical accuracy from the AVTZ basis set, ii) the TC Hamiltonian obtained with the FROGG and that with $\mfrogg$ provide very similar results, iii) the values obtained with $\muuegav$ and $\murscav$ provide very similar results and improve accuracy over the FROGG for the AVDZ basis set, iv) a value of $\mu = 0.5$ provides a good description in the AVDZ basis set while remaining above the exact energy in the AVTZ basis set, which is not the case for $\mu=0.3$. It should also be noticed that when the system becomes strongly correlated (\textit{i.e.} when the bond is stretched), the TC eigenvalues, whatever the Jastrow chosen, do not show any pathological behaviour.

@@

\subsection{Analytical form of $\hat{L}(\bri{1},\bri{2},\bri{3},\mu)$ and computation of related integrals }

@@

 &  \matelem{\phi_i \phi_j \phi_m}{\hat{L}(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} = \\
 +& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} \\
 +& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_2(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} \\
 +& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_3(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} 

@@

 \hat{L}_1(\bri{1},\bri{2},\bri{3},\mu) = w_\mu(\bri{12}) w_\mu(\bri{13}) \bri{12} \cdot \bri{13}

@@

 & \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} = \\

@@

 &  \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n}  \\
 =& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1^x(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} \\  
+ & \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1^y(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} \\  
+ & \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1^z(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n},  

@@

& \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1^x(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} \\  

@@

Therefore, the matrix element of the $\hat{L}_1(\bri{1},\bri{2},\bri{3},\mu)$ is simply 

@@

 & \matelem{\phi_i \phi_j \phi_m}{\hat{L}_1(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} \\

@@

 & \matelem{\phi_i \phi_j \phi_m}{\hat{L}(\bri{1},\bri{2},\bri{3},\mu)}{\phi_k \phi_l \phi_n} \\
